{\bf \huge Wikiplia}:\\ The Free Programming Language\\ That Anyone Can Edit

### 2007-03-20 wikiplia/paper/wikiplia.tex
improve look of main page, but should use templates for this eventually
camera ready

\newcommand\currentrevision{475}
\newcommand\wikipliaurl{http://wikiplia.spacebar.org/}

@@

\subsection{Solution}

Based on these scenarios we conclude that extant social measures such
as revision control conventions are not enough to guarantee freedom.
Even if we think these situations are implausible in the hands of
well-intentioned, well-mannered and capable%
%
\wcite{Year 2000 problem}$^,$%
\wciteshort{Northeast Blackout of 2003}$^,$%
\wcite{Ariane 5 Flight 501}$^,$%
\wcite{Mars Climate Orbiter}$^,$%
\wcite{Mars Polar Lander}$^,$%
\wcite{Mars Rover}$^,$%
\wcite{Mars Pathfinder}$^,$%
\wcite{Gripen\#Crashes}$^,$%
\wcite{missingno.}$^,$%
\wcite{XSS}$^,$%
\wcite{Buffer overflow}$^,$%
\wcite{Therac-25}$^,$%
\wcite{Lothar (storm)}$^,$%
\wcite{Mariner 1}$^,$%
\wcite{Code Red worm}$^,$%
\wcite{SQLSlammer}$^,$%
\wcite{Sandstorm (vehicle)}$^,$%
\wcite{Samy (XSS)}$^,$%
\wcite{Pentium FDIV bug}$^,$%
\wcite{MIM-103 Patriot}$^,$%
\wcite{Windows 95}$^,$%
\wcite{Morris (computer worm)}$^,$%
\wcite{USS Yorktown (CG-48)}
%
software engineers, we wish for our software to remain free even when
in the hands of nefarious and crafty factions who would seek to
fracture\wcite{Fork (software development)} our free software
community. We therefore need a technological and legal solution that
forces the entire development history to be available. Keeping with
Freedom~\fusa, we call this technology and its license {\em Total
Information Awareness} after the successful project of the US
Government with the same name.\z\wcite{Information Awareness Office}

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large
numbers\footnote{The implementation also supports revision numbers of
arbitrary size.} of revisions with small edits (since every change is
saved), so the implementation is engineered to make storing small
revisions very cheap. Particularly, revisions are aggressively
compressed by only storing the newest revision directly, and then a
series of difference ``plans'' that describe how to get the next older
revision from the current one. We compute optimal plans using an
efficient minimal edit distance\wcite{Levenshtein distance}
calculation at the token\wcite{Token (parser)} level. As of
revision~\currentrevision, the database is only 236
kilobytes.\z\wcite{Kilobyte}

@@

% XXX main emergency revert

@@

All popular modern languages are defined via a definitional
interpreter\wcite{JavaScript}$^,$\wcite{Objective
Caml}$^,$\wcite{Perl} with accompanying O'Reilly ``animal''
book.\z\wcite{O'Reilly Media} The work on Wikiplia is unrelated: We
have no animal mascot\wcite{ORLY owl} and the languages are described
by a tower of source-to-source translations on top of a universally
parseable semantic document in XML\wciteshort{Category:ML programming
language family} form.

The author doesn't think\wcite{Thought} much of
musicals,\z\wcite{Musical theatre} to be perfectly\wcite{Thomas
Aquinas} honest,\z\wcite{Honesty} so those are basically\wcite{BASIC}
a no-go. He also feels that the metric system\wcite{Metric system} but
paradoxically\wcite{Pardadox} also time zones\wcite{Time zone} are
pretty overrated. Ketchup\wcite{Ketchup} on eggs\wcite{Egg (food)} is
gross,\z\wcite{Gross} but not quite as gross as foie gras\wcite{Foie
gras}, which more or less has the word\wcite{Word} ``gross'' in its
name\wcite{Name} so duh.\z\wcite{Duh}

@@

We have described Wikiplia, the free programming language that anyone
can edit. Unlike other programming languages, it is designed to
support a variety of freedoms (\ftinker, \fexpression, \fbeer, \fusa,
\ffree, \ftobe) and is explicitly scalable to new freedoms. The system
is implemented in a minimal bootstrapping core, and then built up to
a featured system using its own faculties for extension. XXX more...

>>>>

\newcommand\currentrevision{532}
\newcommand\currentsizekb{250}
\newcommand\wikipliaurl{http://wikiplia.spacebar.org:2222/}

@@

% so that visual gag on ``capable'' doesn't appear at top of page
\vspace{10em}
\hyphenation{capable}
\subsection{Solution}

Based on these scenarios we conclude that extant social measures such
as revision control conventions are not enough to guarantee freedom
in all circumstances.
Even if we think these situations are implausible in the hands of
well-intentioned, well-mannered and capable%
%
\wcite{Year 2000 problem}$^,$%
\wciteshort{Northeast Blackout of 2003}$^,$%
\wcite{Ariane 5 Flight 501}$^,$%
\wcite{Mars Climate Orbiter}$^,$%
\wcite{Mars Polar Lander}$^,$%
\wcite{Mars Rover}$^,$%
\wcite{Mars Pathfinder}$^,$%
\wcite{Gripen\#Crashes}$^,$%
\wcite{missingno.}$^,$%
\wcite{XSS}$^,$%
\wcite{Buffer overflow}$^,$%
\wcite{Therac-25}$^,$%
\wcite{Lothar (storm)}$^,$%
\wcite{Mariner 1}$^,$%
\wcite{Code Red worm}$^,$%
\wcite{SQLSlammer}$^,$%
\wcite{Sandstorm (vehicle)}$^,$%
\wcite{Samy (XSS)}$^,$%
\wcite{Pentium FDIV bug}$^,$%
\wcite{MIM-103 Patriot}$^,$%
\wcite{Windows 95}$^,$%
\wcite{Morris (computer worm)}$^,$%
\wcite{USS Yorktown (CG-48)}
%
software engineers, we wish for our software to remain free even when
in the hands of nefarious and crafty factions who would seek to
fracture\wcite{Fork (software development)} our free software
community. We therefore need a technological and legal solution that
forces the entire development history to be available. Keeping with
Freedom~\fusa, we call this technology and its license {\em Total
Information Awareness} after the successful project of the US
Government with the same name.\z\wcite{Information Awareness Office}

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large
numbers\footnote{The implementation also supports revision numbers of
arbitrary size.} of revisions with small edits (since every change is
saved), so the implementation is engineered to make storing small
revisions very cheap. Particularly, revisions are aggressively
compressed by only storing the newest revision directly, and then a
series of difference ``plans'' that describe how to get the next older
revision from the current one. We compute optimal plans using an
efficient minimal edit distance\wcite{Levenshtein distance}
calculation at the token\wcite{Token (parser)} level. As of
revision~\currentrevision, the database is only \currentsizekb\ 
kilobytes.\z\wcite{Kilobyte}

@@

It is very easy to make mistakes that render the system unusable.
Therefore Wikiplia supports the ability to safely revert to a previous
version of a key. When making changes to the {\tt main} key this
ability can be accidentally disabled, so the complex functionality of
{\tt main} was split off to a new key called {\tt main-go} at revision
9; the {\tt main} key now only dispatches to {\tt main-go} but
provides an {\tt emergency-revert} action that automatiaclly reverts
{\tt main-go} to its previous revision in case it is damaged and the
site is unusable.

@@

All popular modern languages are defined via a definitional
interpreter\wcite{JavaScript}$^,$\wcite{Objective
Caml}$^,$\wcite{Perl} with accompanying O'Reilly ``animal''
book.\z\wcite{O'Reilly Media} The work on Wikiplia is unrelated: We
have no animal mascot\wcite{ORLY owl} and the languages are described
by a tower of source-to-source translations\wcite{Translation} on top
of a universally parseable semantic document in
XML\wciteshort{Category:ML programming language family} form.

The author\wcite{Author} doesn't think\wcite{Thought} much of
musicals,\z\wcite{Musical theatre} to be perfectly\wcite{Thomas
Aquinas} honest,\z\wcite{Honesty} so those are basically\wcite{BASIC}
a no-go. He also feels that the metric system\wcite{Metric system} but
paradoxically\wcite{Pardadox} also time zones\wcite{Time zone} are
pretty overrated. Ketchup\wcite{Ketchup} on eggs\wcite{Egg (food)} is
gross,\z\wcite{Gross} but not quite as gross as foie gras\wcite{Foie
gras}, which more or less has the word\wcite{Word} ``gross'' in its
name\wcite{Name} so duh.\z\wcite{Duh}

@@

We have described Wikiplia, the free programming language that anyone
can edit. Unlike other programming languages, it is designed to
support a variety of freedoms (\ftinker, \fexpression, \fbeer, \fusa,
\ffree, \ftobe) and is explicitly scalable to new freedoms. Wikiplia
is implemented in a minimal bootstrapping core based on freedom-aware
technologies such as XML, and then built up to a featured system using
its own faculties for extension. However, much work remains to be
done. We invite you to join us!
